---
title: "Analyzing Europes Top 5 European Soccer Leagues: 09/10 - 18/19"
output: html_document
date: "2023-02-01"
---
This document is meant to analyze soccer data in Europes Top 5 European League from the 09/10 season to the 18/19 season. The idea is touch on the competitiveness of each league. What does competitive actually mean? 

Competitiveness is a vague term - we will look at winners, top 4 contenders and points earned from the bottom teams from the top teams to compare whether one league really shines or dulls compared to the others. 

This analysis is conducted using a combination of R programming & SQL. Please view the code file if you want a more in-depth look at my code. This markdown file will only present final outputs such as tables & graphs.  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#read in files
#Packages
library(lubridate)
library(tidyverse)
library(dplyr)
library(sqldf)

#Read in Data + Clean Data
prem0910 = read.csv("input/Premier League/prem0910.csv")
prem1011 = read.csv("input/Premier League/prem1011.csv")
prem1112 = read.csv("input/Premier League/prem1112.csv")
prem1213 = read.csv("input/Premier League/prem1213.csv")
prem1314 = read.csv("input/Premier League/prem1314.csv")
prem1415 = read.csv("input/Premier League/prem1415.csv")
prem1516 = read.csv("input/Premier League/prem1516.csv")
prem1617 = read.csv("input/Premier League/prem1617.csv")
prem1718 = read.csv("input/Premier League/prem1718.csv")
prem1819 = read.csv("input/Premier League/prem1819.csv")

#View(prem0910) #YYYY/MM/DD
#View(prem1415) #DD/MM/yy
#View(prem1819) #DD/MM/YYYY

#CONDENSE DATAFRAMES TO SPECIFIC COLUMNS
prem0910 = prem0910[,c(1:23)]
prem1011 = prem1011[,c(1:23)]
prem1112 = prem1112[,c(1:23)]
prem1213 = prem1213[,c(1:23)]
prem1314 = prem1314[,c(1:23)]
prem1415 = prem1415[,c(1:23)]
prem1516 = prem1516[,c(1:23)]
prem1617 = prem1617[,c(1:23)]
prem1718 = prem1718[,c(1:23)]
prem1819 = prem1819[,c(1:23)]

#MERGE ALL SEASON DATA TOGETHER
premalldata = do.call("rbind", list(prem0910, prem1011, prem1112, prem1213, prem1314, prem1415,
                                    prem1516, prem1617, prem1718, prem1819))
#View(premalldata)

#FIX DATE

premalldata$Date = parse_date_time(premalldata$Date,orders = c("dmy","ymd"))

#ADD SEASON
premalldata$season[1:380] = "09/10"
premalldata$season[381:760] = "10/11"
premalldata$season[761:1140] = "11/12"
premalldata$season[1141:1520] = "12/13"
premalldata$season[1521:1900] = "13/14"
premalldata$season[1901:2280] = "14/15"
premalldata$season[2281:2660] = "15/16"
premalldata$season[2661:3040] = "16/17"
premalldata$season[3041:3420] = "17/18"
premalldata$season[3421:3800] = "18/19"

#ADD POINTS
premalldata$HomeTeamPoints = ifelse(premalldata$FTHG > premalldata$FTAG, 3,
                                    ifelse(premalldata$FTHG == premalldata$FTAG,1,0))
premalldata$AwayTeamPoints = ifelse(premalldata$FTHG > premalldata$FTAG, 0,
                                    ifelse(premalldata$FTHG == premalldata$FTAG,1,3))

#GET RID OF UNNECCESSARY COLUMNS
premalldata = premalldata[,c(1:6,24:26)]

#REORDER COLUMNS
premalldata = premalldata[,c(1,2,7,3,4,5,6,8,9)]

#-------------------------------------------------------------------------------------------------
#Bundesliga

bund0910 = read.csv("input/Bundesliga/Bundesliga0910.csv")
bund1011 = read.csv("input/Bundesliga/Bundesliga1011.csv")
bund1112 = read.csv("input/Bundesliga/Bundesliga1112.csv")
bund1213 = read.csv("input/Bundesliga/Bundesliga1213.csv")
bund1314 = read.csv("input/Bundesliga/Bundesliga1314.csv")
bund1415 = read.csv("input/Bundesliga/Bundesliga1415.csv")
bund1516 = read.csv("input/Bundesliga/Bundesliga1516.csv")
bund1617 = read.csv("input/Bundesliga/Bundesliga1617.csv")
bund1718 = read.csv("input/Bundesliga/Bundesliga1718.csv")
bund1819 = read.csv("input/Bundesliga/Bundesliga1819.csv")

#think its the same data format as prem
#View(bund0910)
#View(bund1314)
#View(bund1819)

#CONDENSE DATAFRAMES TO SPECIFIC COLUMNS
bund0910 = bund0910[,c(1:23)]
bund1011 = bund1011[,c(1:23)]
bund1112 = bund1112[,c(1:23)]
bund1213 = bund1213[,c(1:23)]
bund1314 = bund1314[,c(1:23)]
bund1415 = bund1415[,c(1:23)]
bund1516 = bund1516[,c(1:23)]
bund1617 = bund1617[,c(1:23)]
bund1718 = bund1718[,c(1:23)]
bund1819 = bund1819[,c(1:23)]

#MERGE ALL SEASON DATA TOGETHER
bundalldata = do.call("rbind", list(bund0910, bund1011, bund1112, bund1213, bund1314, bund1415,
                                    bund1516, bund1617, bund1718, bund1819))
#View(bundalldata)

#FIX DATE

bundalldata$Date = parse_date_time(bundalldata$Date,orders = c("dmy","ymd"))

#ADD SEASON
bundalldata$season[1:306] = "09/10"
bundalldata$season[307:612] = "10/11"
bundalldata$season[613:918] = "11/12"
bundalldata$season[919:1224] = "12/13"
bundalldata$season[1225:1530] = "13/14"
bundalldata$season[1531:1836] = "14/15"
bundalldata$season[1837:2142] = "15/16"
bundalldata$season[2143:2448] = "16/17"
bundalldata$season[2449:2754] = "17/18"
bundalldata$season[2755:3060] = "18/19"

bundalldata$B365H = NULL

#ADD POINTS
bundalldata$HomeTeamPoints = ifelse(bundalldata$FTHG > bundalldata$FTAG, 3,
                                    ifelse(bundalldata$FTHG == bundalldata$FTAG,1,0))
bundalldata$AwayTeamPoints = ifelse(bundalldata$FTHG > bundalldata$FTAG, 0,
                                    ifelse(bundalldata$FTHG == bundalldata$FTAG,1,3))

#GET RID OF UNNECCESSARY COLUMNS
bundalldata = bundalldata[,c(1:6,23:25)]

#REORDER COLUMNS
bundalldata = bundalldata[,c(1,2,7,3,4,5,6,8,9)]

#-----------------------------------------------------------------------------------------------------

#Serie A

italy0910 = read.csv("input/Serie A/SerieA0910.csv")
italy1011 = read.csv("input/Serie A/SerieA1011.csv")
italy1112 = read.csv("input/Serie A/SerieA1112.csv")
italy1213 = read.csv("input/Serie A/SerieA1213.csv")
italy1314 = read.csv("input/Serie A/SerieA1314.csv")
italy1415 = read.csv("input/Serie A/SerieA1415.csv")
italy1516 = read.csv("input/Serie A/SerieA1516.csv")
italy1617 = read.csv("input/Serie A/SerieA1617.csv")
italy1718 = read.csv("input/Serie A/SerieA1718.csv")
italy1819 = read.csv("input/Serie A/SerieA1819.csv")

#View(italy0910)
#View(italy1415)
#View(italy1819)
#again dates think are same as prem

#CONDENSE DATAFRAMES TO SPECIFIC COLUMNS
italy0910 = italy0910[,c(1:23)]
italy1011 = italy1011[,c(1:23)]
italy1112 = italy1112[,c(1:23)]
italy1213 = italy1213[,c(1:23)]
italy1314 = italy1314[,c(1:23)]
italy1415 = italy1415[,c(1:23)]
italy1516 = italy1516[,c(1:23)]
italy1617 = italy1617[,c(1:23)]
italy1718 = italy1718[,c(1:23)]
italy1819 = italy1819[,c(1:23)]

#MERGE ALL SEASON DATA TOGETHER
italyalldata = do.call("rbind", list(italy0910, italy1011, italy1112, italy1213, italy1314, italy1415,
                                     italy1516, italy1617, italy1718, italy1819))
#View(italyalldata)

italyalldata$B365H = NULL

#FIX DATE

italyalldata$Date = parse_date_time(italyalldata$Date,orders = c("dmy","ymd"))

#ADD SEASON
italyalldata$season[1:380] = "09/10"
italyalldata$season[381:760] = "10/11"
italyalldata$season[761:1140] = "11/12"
italyalldata$season[1141:1520] = "12/13"
italyalldata$season[1521:1900] = "13/14"
italyalldata$season[1901:2280] = "14/15"
italyalldata$season[2281:2660] = "15/16"
italyalldata$season[2661:3040] = "16/17"
italyalldata$season[3041:3420] = "17/18"
italyalldata$season[3421:3800] = "18/19"

#ADD POINTS
italyalldata$HomeTeamPoints = ifelse(italyalldata$FTHG > italyalldata$FTAG, 3,
                                     ifelse(italyalldata$FTHG == italyalldata$FTAG,1,0))
italyalldata$AwayTeamPoints = ifelse(italyalldata$FTHG > italyalldata$FTAG, 0,
                                     ifelse(italyalldata$FTHG == italyalldata$FTAG,1,3))

#GET RID OF UNNECCESSARY COLUMNS
italyalldata = italyalldata[,c(1:6,23:25)]

#REORDER COLUMNS
italyalldata = italyalldata[,c(1,2,7,3,4,5,6,8,9)]

#--------------------------------------------------------------------------------------------------

#La Liga

laliga0910 = read.csv("input/La Liga/laliga0910.csv")
laliga1011 = read.csv("input/La Liga/laliga1011.csv")
laliga1112 = read.csv("input/La Liga/laliga1112.csv")
laliga1213 = read.csv("input/La Liga/laliga1213.csv")
laliga1314 = read.csv("input/La Liga/laliga1314.csv")
laliga1415 = read.csv("input/La Liga/laliga1415.csv")
laliga1516 = read.csv("input/La Liga/laliga1516.csv")
laliga1617 = read.csv("input/La Liga/laliga1617.csv")
laliga1718 = read.csv("input/La Liga/laliga1718.csv")
laliga1819 = read.csv("input/La Liga/laliga1819.csv")

#View(laliga0910)
#View(laliga1415)
#View(laliga1819)
#again dates think are same as prem

#CONDENSE DATAFRAMES TO SPECIFIC COLUMNS
laliga0910 = laliga0910[,c(1:23)]
laliga1011 = laliga1011[,c(1:23)]
laliga1112 = laliga1112[,c(1:23)]
laliga1213 = laliga1213[,c(1:23)]
laliga1314 = laliga1314[,c(1:23)]
laliga1415 = laliga1415[,c(1:23)]
laliga1516 = laliga1516[,c(1:23)]
laliga1617 = laliga1617[,c(1:23)]
laliga1718 = laliga1718[,c(1:23)]
laliga1819 = laliga1819[,c(1:23)]

#MERGE ALL SEASON DATA TOGETHER
laligaalldata = do.call("rbind", list(laliga0910, laliga1011, laliga1112, laliga1213, laliga1314, laliga1415,
                                      laliga1516, laliga1617, laliga1718, laliga1819))
#View(laligaalldata)

laligaalldata$B365H = NULL

#FIX DATE

laligaalldata$Date = parse_date_time(laligaalldata$Date,orders = c("dmy","ymd"))

#ADD SEASON
laligaalldata$season[1:380] = "09/10"
laligaalldata$season[381:760] = "10/11"
laligaalldata$season[761:1140] = "11/12"
laligaalldata$season[1141:1520] = "12/13"
laligaalldata$season[1521:1900] = "13/14"
laligaalldata$season[1901:2280] = "14/15"
laligaalldata$season[2281:2660] = "15/16"
laligaalldata$season[2661:3040] = "16/17"
laligaalldata$season[3041:3420] = "17/18"
laligaalldata$season[3421:3800] = "18/19"

#ADD POINTS
laligaalldata$HomeTeamPoints = ifelse(laligaalldata$FTHG > laligaalldata$FTAG, 3,
                                      ifelse(laligaalldata$FTHG == laligaalldata$FTAG,1,0))
laligaalldata$AwayTeamPoints = ifelse(laligaalldata$FTHG > laligaalldata$FTAG, 0,
                                      ifelse(laligaalldata$FTHG == laligaalldata$FTAG,1,3))

#GET RID OF UNNECCESSARY COLUMNS
laligaalldata = laligaalldata[,c(1:6,23:25)]

#REORDER COLUMNS
laligaalldata = laligaalldata[,c(1,2,7,3,4,5,6,8,9)]

#-------------------------------------------------------------------------------------------------------

#Ligue 1

french0910 = read.csv("input/Ligue 1/French0910.csv")
french1011 = read.csv("input/Ligue 1/French1011.csv")
french1112 = read.csv("input/Ligue 1/French1112.csv")
french1213 = read.csv("input/Ligue 1/French1213.csv")
french1314 = read.csv("input/Ligue 1/French1314.csv")
french1415 = read.csv("input/Ligue 1/French1415.csv")
french1516 = read.csv("input/Ligue 1/French1516.csv")
french1617 = read.csv("input/Ligue 1/French1617.csv")
french1718 = read.csv("input/Ligue 1/French1718.csv")
french1819 = read.csv("input/Ligue 1/French1819.csv")

#View(french0910)
#View(french1415)
#View(french1819)
#again dates think are same as prem

#CONDENSE DATAFRAMES TO SPECIFIC COLUMNS
french0910 = french0910[,c(1:23)]
french1011 = french1011[,c(1:23)]
french1112 = french1112[,c(1:23)]
french1213 = french1213[,c(1:23)]
french1314 = french1314[,c(1:23)]
french1415 = french1415[,c(1:23)]
french1516 = french1516[,c(1:23)]
french1617 = french1617[,c(1:23)]
french1718 = french1718[,c(1:23)]
french1819 = french1819[,c(1:23)]

#MERGE ALL SEASON DATA TOGETHER
frenchalldata = do.call("rbind", list(french0910, french1011, french1112, french1213, french1314, french1415,
                                      french1516, french1617, french1718, french1819))
#View(frenchalldata)

frenchalldata$B365H = NULL

#FIX DATE

frenchalldata$Date = parse_date_time(frenchalldata$Date,orders = c("dmy","ymd"))

#ADD SEASON
frenchalldata$season[1:380] = "09/10"
frenchalldata$season[381:760] = "10/11"
frenchalldata$season[761:1140] = "11/12"
frenchalldata$season[1141:1520] = "12/13"
frenchalldata$season[1521:1900] = "13/14"
frenchalldata$season[1901:2280] = "14/15"
frenchalldata$season[2281:2660] = "15/16"
frenchalldata$season[2661:3040] = "16/17"
frenchalldata$season[3041:3420] = "17/18"
frenchalldata$season[3421:3800] = "18/19"

#ADD POINTS
frenchalldata$HomeTeamPoints = ifelse(frenchalldata$FTHG > frenchalldata$FTAG, 3,
                                      ifelse(frenchalldata$FTHG == frenchalldata$FTAG,1,0))
frenchalldata$AwayTeamPoints = ifelse(frenchalldata$FTHG > frenchalldata$FTAG, 0,
                                      ifelse(frenchalldata$FTHG == frenchalldata$FTAG,1,3))

#GET RID OF UNNECCESSARY COLUMNS
frenchalldata = frenchalldata[,c(1:6,23:25)]

#REORDER COLUMNS
frenchalldata = frenchalldata[,c(1,2,7,3,4,5,6,8,9)]
```

# Let's Start with the Premier League

## Premier League Tables 09/10 - 18/19 {.tabset}

```{r analysis PL, echo = FALSE, results='asis'}

for(t in unique(premalldata$season)) {
  cat("###", t,'\n') #open each tab and label it
  plots <- 
    premalldata%>%
    filter(season == t)%>%
    select(Team = HomeTeam,Points = HomeTeamPoints) %>%
    bind_rows(premalldata%>%
              filter(season == t)%>%
              select(Team = AwayTeam,Points = AwayTeamPoints))%>%
    group_by(Team)%>%
    summarise(Points = sum(Points))%>%
    arrange(desc(Points))
  print(knitr::kable(plots))
  cat('\n\n') #closse each tab
    
}


```

## Premier League Winner's & Top 4 09/10 - 18/19

```{r PL Winners, echo=FALSE, comment=NA}

premtable <- as.data.frame(sqldf("SELECT Team, sum(Points) as Points, season
          FROM 
          (SELECT HomeTeam as Team, sum(HomeTeamPoints) as Points, season FROM premalldata GROUP BY Team, season
          UNION
          SELECT AwayTeam as Team, sum(AwayTeamPoints) as Points, season FROM premalldata GROUP BY Team,season)
      GROUP BY Team, season
      ORDER BY season, Points desc"))

premtable$Rank <-rep(c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20),times=10)


#Add the ranks to premalldata for home and awayteams

premalldata <- sqldf("SELECT Div, Date, p1.season, HomeTeam, AwayTeam, FTHG, FTAG, HomeTeamPoints, AwayTeamPoints, p2.Rank as HomeRank, p3.Rank as AwayRank
      FROM premalldata p1
      LEFT JOIN premtable p2 ON p1.HomeTeam = p2.Team AND p1.season = p2.season
      LEFT JOIN premtable p3 ON p1.AwayTeam = p3.Team AND p1.season = p3.season")

#SQL
prem_champs <- sqldf("SELECT Team as Champion, season
      FROM premtable 
      WHERE Rank = 1")

print(prem_champs)

premtop4 <- sqldf("SELECT Team, Count(*) as Top4Finishes
      FROM premtable 
      WHERE Rank BETWEEN 1 AND 4
      GROUP BY Team
      ORDER by Top4Finishes DESC")
print(premtop4)

```

This is why people enjoy watching the PL. 10 seasons, no team was ever champion twice in a row during this time. However, in the context of the top 4, this is largely fought between 5 teams during this period. This may indicate more of a gap between the top 4-5 teams and the mid table sides.

# Let's move on to the German Bundesliga

## Bundesliga Tables 09/10 - 18/19 {.tabset}

```{r analysis Bundes, echo = FALSE, results='asis'}

for(t in unique(bundalldata$season)) {
  cat("###", t,'\n') #open each tab and label it
  plots <- 
    bundalldata%>%
    filter(season == t)%>%
    select(Team = HomeTeam,Points = HomeTeamPoints) %>%
    bind_rows(bundalldata%>%
              filter(season == t)%>%
              select(Team = AwayTeam,Points = AwayTeamPoints))%>%
    group_by(Team)%>%
    summarise(Points = sum(Points))%>%
    arrange(desc(Points))
  print(knitr::kable(plots))
  cat('\n\n') #closse each tab
    
}

```

## Bundesliga Winner's & Top 4 09/10 - 18/19

```{r Bundes Winners, echo=FALSE, comment = NA}

bundtable <- as.data.frame(sqldf("SELECT Team, sum(Points) as Points, season
          FROM 
          (SELECT HomeTeam as Team, sum(HomeTeamPoints) as Points, season FROM bundalldata GROUP BY Team, season
          UNION
          SELECT AwayTeam as Team, sum(AwayTeamPoints) as Points, season FROM bundalldata GROUP BY Team,season)
      GROUP BY Team, season
      ORDER BY season, Points desc"))

bundtable$Rank <-rep(c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18),times=10)


#Add the ranks to bundalldata for home and awayteams

bundalldata <- sqldf("SELECT Div, Date, p1.season, HomeTeam, AwayTeam, FTHG, FTAG, HomeTeamPoints, AwayTeamPoints, p2.Rank as HomeRank, p3.Rank as AwayRank
      FROM bundalldata p1
      LEFT JOIN bundtable p2 ON p1.HomeTeam = p2.Team AND p1.season = p2.season
      LEFT JOIN bundtable p3 ON p1.AwayTeam = p3.Team AND p1.season = p3.season")

#SQL
bund_champs <- sqldf("SELECT Team as Champion, season
      FROM bundtable 
      WHERE Rank = 1")

print(bund_champs)

bundtop4 <- sqldf("SELECT Team, Count(*) as Top4Finishes
      FROM bundtable 
      WHERE Rank BETWEEN 1 AND 4
      GROUP BY Team
      ORDER by Top4Finishes DESC")
print(bundtop4)

```

This is why the Bundesliga comes under much scrutiny. Bayern Munich dominates this league and was champions 8/10 times during this period. The top 4 is a bit more interesting though. You see a few more teams getting into these positions than the premier league, identifying that after Bayern, there is a bit more balance to the 2nd through mid table positions. 


# Italian Serie A

## Serie A Tables 09/10 - 18/19 {.tabset}

```{r analysis Serie A, echo = FALSE, results='asis'}

for(t in unique(italyalldata$season)) {
  cat("###", t,'\n') #open each tab and label it
  plots <- 
    italyalldata%>%
    filter(season == t)%>%
    select(Team = HomeTeam,Points = HomeTeamPoints) %>%
    bind_rows(italyalldata%>%
              filter(season == t)%>%
              select(Team = AwayTeam,Points = AwayTeamPoints))%>%
    group_by(Team)%>%
    summarise(Points = sum(Points))%>%
    arrange(desc(Points))
  print(knitr::kable(plots))
  cat('\n\n') #closse each tab
    
}

```

## Serie A Winner's & Top 4 09/10 - 18/19

```{r Winners Serie A, echo=FALSE, comment = NA}

italytable <- as.data.frame(sqldf("SELECT Team, sum(Points) as Points, season
          FROM 
          (SELECT HomeTeam as Team, sum(HomeTeamPoints) as Points, season FROM italyalldata GROUP BY Team, season
          UNION
          SELECT AwayTeam as Team, sum(AwayTeamPoints) as Points, season FROM italyalldata GROUP BY Team,season)
      GROUP BY Team, season
      ORDER BY season, Points desc"))

italytable$Rank <-rep(c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20),times=10)


#Add the ranks to italyalldata for home and awayteams

italyalldata <- sqldf("SELECT Div, Date, p1.season, HomeTeam, AwayTeam, FTHG, FTAG, HomeTeamPoints, AwayTeamPoints, p2.Rank as HomeRank, p3.Rank as AwayRank
      FROM italyalldata p1
      LEFT JOIN italytable p2 ON p1.HomeTeam = p2.Team AND p1.season = p2.season
      LEFT JOIN italytable p3 ON p1.AwayTeam = p3.Team AND p1.season = p3.season")

#SQL
italy_champs <- sqldf("SELECT Team as Champion, season
      FROM italytable 
      WHERE Rank = 1")

print(italy_champs)

italytop4 <- sqldf("SELECT Team, Count(*) as Top4Finishes
      FROM italytable 
      WHERE Rank BETWEEN 1 AND 4
      GROUP BY Team
      ORDER by Top4Finishes DESC")
print(italytop4)

```

The Serie A followed a similar pattern to the Bundesliga during this period. Juventus were champions 8/10 times, 8 times in a row actually. Fans outside of Juventus (and Bayern) do not want to see this. They want to watch competition, new winners - it's more interesting. Thankfully the top 4 battle gives that interest. Again, similar to the Bundesliga we see many more teams than the PL winning these positions during this time, indicating that the gap from 2nd to mid table is often quite competitive. I would say there's even a bit more balance here than the Bundesliga.  

# Spain's La Liga

## La Liga Tables 09/10 - 18/19 {.tabset}

```{r analysis La Liga, echo = FALSE, results='asis'}

for(t in unique(laligaalldata$season)) {
  cat("###", t,'\n') #open each tab and label it
  plots <- 
    laligaalldata%>%
    filter(season == t)%>%
    select(Team = HomeTeam,Points = HomeTeamPoints) %>%
    bind_rows(laligaalldata%>%
              filter(season == t)%>%
              select(Team = AwayTeam,Points = AwayTeamPoints))%>%
    group_by(Team)%>%
    summarise(Points = sum(Points))%>%
    arrange(desc(Points))
  print(knitr::kable(plots))
  cat('\n\n') #closse each tab
    
}

```

## La Liga Winner's & Top 4 09/10 - 18/19

```{r Winners La Liga, echo=FALSE, comment=NA}

laligatable <- as.data.frame(sqldf("SELECT Team, sum(Points) as Points, season
          FROM 
          (SELECT HomeTeam as Team, sum(HomeTeamPoints) as Points, season FROM laligaalldata GROUP BY Team, season
          UNION
          SELECT AwayTeam as Team, sum(AwayTeamPoints) as Points, season FROM laligaalldata GROUP BY Team,season)
      GROUP BY Team, season
      ORDER BY season, Points desc"))

laligatable$Rank <-rep(c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20),times=10)


#Add the ranks to laligaalldata for home and awayteams

laligaalldata <- sqldf("SELECT Div, Date, p1.season, HomeTeam, AwayTeam, FTHG, FTAG, HomeTeamPoints, AwayTeamPoints, p2.Rank as HomeRank, p3.Rank as AwayRank
      FROM laligaalldata p1
      LEFT JOIN laligatable p2 ON p1.HomeTeam = p2.Team AND p1.season = p2.season
      LEFT JOIN laligatable p3 ON p1.AwayTeam = p3.Team AND p1.season = p3.season")

#SQL
laliga_champs <- sqldf("SELECT Team as Champion, season
      FROM laligatable 
      WHERE Rank = 1")

print(laliga_champs)

laligatop4 <- sqldf("SELECT Team, Count(*) as Top4Finishes
      FROM laligatable 
      WHERE Rank BETWEEN 1 AND 4
      GROUP BY Team
      ORDER by Top4Finishes DESC")
print(laligatop4)

```

Ah La Liga. During this period, it was the peak of European Football. Real Madrid vs Barcelona - Ronaldo vs Messi - were the most anticipated games on the calendar (arguably they still are). But as many other domestic league fanbases point out, it is a league largely dominated by 2 with no room for the rest. We see Barca and RM winning 9/10 titles during this time, the top 4 is dominated by 4 teams over this period as well. Once Messi and Ronaldo left, the hype died down quite a bit and many eyes have turned from La Liga to PL as it's deemed a more competitve and watchable league. 

# France's Ligue 1

## Ligue 1 Tables 09/10 - 18/19 {.tabset}

```{r analysis Ligue 1, echo = FALSE, results='asis'}

for(t in unique(frenchalldata$season)) {
  cat("###", t,'\n') #open each tab and label it
  plots <- 
    frenchalldata%>%
    filter(season == t)%>%
    select(Team = HomeTeam,Points = HomeTeamPoints) %>%
    bind_rows(frenchalldata%>%
              filter(season == t)%>%
              select(Team = AwayTeam,Points = AwayTeamPoints))%>%
    group_by(Team)%>%
    summarise(Points = sum(Points))%>%
    arrange(desc(Points))
  print(knitr::kable(plots))
  cat('\n\n') #closse each tab
    
}

```

## Ligue 1 Winner's & Top 4 09/10 - 18/19

```{r Winners Ligue 1, echo=FALSE,comment=NA}

frenchtable <- as.data.frame(sqldf("SELECT Team, sum(Points) as Points, season
          FROM 
          (SELECT HomeTeam as Team, sum(HomeTeamPoints) as Points, season FROM frenchalldata GROUP BY Team, season
          UNION
          SELECT AwayTeam as Team, sum(AwayTeamPoints) as Points, season FROM frenchalldata GROUP BY Team,season)
      GROUP BY Team, season
      ORDER BY season, Points desc"))

frenchtable$Rank <-rep(c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20),times=10)


#Add the ranks to frenchalldata for home and awayteams

frenchalldata <- sqldf("SELECT Div, Date, p1.season, HomeTeam, AwayTeam, FTHG, FTAG, HomeTeamPoints, AwayTeamPoints, p2.Rank as HomeRank, p3.Rank as AwayRank
      FROM frenchalldata p1
      LEFT JOIN frenchtable p2 ON p1.HomeTeam = p2.Team AND p1.season = p2.season
      LEFT JOIN frenchtable p3 ON p1.AwayTeam = p3.Team AND p1.season = p3.season")

#SQL
french_champs <- sqldf("SELECT Team as Champion, season
      FROM frenchtable 
      WHERE Rank = 1")

print(french_champs)

frenchtop4 <- sqldf("SELECT Team, Count(*) as Top4Finishes
      FROM frenchtable 
      WHERE Rank BETWEEN 1 AND 4
      GROUP BY Team
      ORDER by Top4Finishes DESC")
print(frenchtop4)

```

Ligue 1 is difficult to make sense of during this period. From the data we see 6 different champions in 10 seasons, the most of any of the top 5 leagues. We see a top 4 largely contested by 5 teams, but others tend to sneak in as well some years. So why is this league so largely criticized? PSG were bought in 2011 and you can see starting from the 12/13 season, they won 6/7 seasons (15/16 should be PSG, not Lyon). Often they win by large margins too. People don't see the league as competitive anymore due to their dominance. We will explore a bit in the next section how competitive the league as a whole is. 

# Points Earned from the Bottom 6 Teams vs Top 6 Teams per Season

This is such a random measure I thought to look into to compare competitiveness in each league. I often here from commentary that anyone can beat anyone on any given day in England and that's why their league is the best in the world. Apparently these results do not happen anywhere else? Let's take a look. 

```{r, Points Earned Bottom 6 vs Top 6, echo=FALSE, results='asis'}

prembottom6 <- sqldf("SELECT season, sum(CASE 
      WHEN HomeRank BETWEEN 1 and 6 and AwayRank BETWEEN 15 and 20 THEN AwayTeamPoints
      WHEN HomeRank BETWEEN 15 and 20 and AwayRank BETWEEN 1 and 6 THEN HomeTeamPoints END) AS PremBottom6Points
      FROM premalldata
      GROUP BY season")

bundbottom6 <- sqldf("SELECT season, sum(CASE 
      WHEN HomeRank BETWEEN 1 and 6 and AwayRank BETWEEN 15 and 20 THEN AwayTeamPoints
      WHEN HomeRank BETWEEN 15 and 20 and AwayRank BETWEEN 1 and 6 THEN HomeTeamPoints END) AS BundBottom6Points
      FROM bundalldata
      GROUP BY season")

italybottom6 <-  sqldf("SELECT season, sum(CASE 
      WHEN HomeRank BETWEEN 1 and 6 and AwayRank BETWEEN 15 and 20 THEN AwayTeamPoints
      WHEN HomeRank BETWEEN 15 and 20 and AwayRank BETWEEN 1 and 6 THEN HomeTeamPoints END) AS ItalyBottom6Points
      FROM italyalldata
      GROUP BY season")

laligabottom6 <- sqldf("SELECT season, sum(CASE 
      WHEN HomeRank BETWEEN 1 and 6 and AwayRank BETWEEN 15 and 20 THEN AwayTeamPoints
      WHEN HomeRank BETWEEN 15 and 20 and AwayRank BETWEEN 1 and 6 THEN HomeTeamPoints END) AS LaLigaBottom6Points
      FROM laligaalldata
      GROUP BY season")

frenchbottom6 <-  sqldf("SELECT season, sum(CASE 
      WHEN HomeRank BETWEEN 1 and 6 and AwayRank BETWEEN 15 and 20 THEN AwayTeamPoints
      WHEN HomeRank BETWEEN 15 and 20 and AwayRank BETWEEN 1 and 6 THEN HomeTeamPoints END) AS FrenchBottom6Points
      FROM frenchalldata
      GROUP BY season")

bottom6 <- prembottom6%>%
  left_join(bundbottom6, by = "season")%>%
  left_join(italybottom6, by = "season")%>%
  left_join(laligabottom6, by = "season")%>%
  left_join(frenchbottom6, by = "season")


bottom6 <-  bottom6%>%
  pivot_longer(!season, names_to = "League",
               values_to = "Points")

ggplot(bottom6)+
  geom_line(aes(x = season, y = Points, group = League, color = League))+
  geom_point(aes(x = season, y = Points, group = League, color = League))+
  ggtitle("Points Earned per Season from Bottom 6 Teams vs \n Top 6 Teams 09/10 - 18/19")



```

So what does this tell us? Well, that there isn't much in it.

You might notice the bundesliga on multiple occasions is at the bottom of these rankings and never at the top.

You might notice the french league is on top on multiple occasions and never at the bottom. (15/16 needs to be cautious as for some reason that I cannot figure out, PSG is not appearing as the top team in 15/16 in some calculations so this number is inflated. My guess would be between 12-24 points which would still make it pretty high)

And then you have the Prem, Italian Serie A and Spanish La Liga all hovering in the middle each having a couple occasions where they're on top and a couple occasions where theyr'e on the bottom

Meaning what? Well that soccer is a sport. None of these top sides completely obliterate the bottom sides year after year, nor lose more often to bottom sides year after year. 

The media frenzy that anyone can beat anyone on any given day in the premier league is probably true. But no more true that any of the other leagues. You just hear about a result more in the premier league because it has the most global media attention and marketing


